---
title: "Covid91 vaccine study Final2020"
author: "Laurice Maayah"
date: "Nov 3, 2020"
output: html_document
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

# Females


```{r}
dataF=filter(FinalData,sex=='gal')
```

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
barchartGC(~infected + treatment,data=dataF)
barchartGC(~infected + treatment,data=dataF, type="percent")

```
  
The females who got the vaccine are less likely to get infected than the females who got the placebo.

```{r}
table1 <- xtabs(~infected + treatment,data=dataF)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
The odds ratio 0.5 to 1 means females who got the vaccine are half as likely to get Covid-19 as females who got the placebo. 

# Males 

```{r}
dataM=filter(FinalData,sex=='guy')
```

<!--Divide this section into two sub-sections:  One for your descriptive results and one for your inferential results.!-->

```{r}
barchartGC(~infected + treatment,data=dataM)
barchartGC(~infected + treatment,data=dataM, type="percent")

```

The males who got the vaccine are more likely to get infected than the males who got the placebo.

```{r}
table1 <- xtabs(~infected + treatment,data=dataM)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odds ratio is 1.5 to 1 so that males who have received the vaccine are 1.5 times more likely than males who have received the placebo to be infected with Covid-19, and the 95% confidence interval is greater than 1 so we know the outcome is important.

#  LGBTQ   

```{r}
dataQ=filter(FinalData,LGBTQ=='gay')
```

<!--Divide this section into two sub-sections:  One for your descriptive results and one for your inferential results.!-->

```{r}
barchartGC(~infected + treatment,data=dataQ)
barchartGC(~infected + treatment,data=dataQ, type="percent")

```

People in the LGBTQ community who obtained the vaccine were more likely than those who received placebo to get Covid 19.

```{r}
table1 <- xtabs(~infected + treatment,data=dataQ)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
The odd ratio is 1.5 to 1 means people in the LGBTQ community are 1.5 times more likely to get infected with Covid-19. the 95% confidence interval is greater than 1 so we know the outcome is important. 


# Drugusers
```{r}
dataD=filter(FinalData,DrugUser=='yes')
```

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

```{r}
barchartGC(~infected + treatment,data=dataD)
barchartGC(~infected + treatment,data=dataD, type="percent")

```

The people who got the vaccine are less likely to get Covid-19 than the people who got the placebo.

```{r}
table1 <- xtabs(~infected + treatment,data=dataD)
rowPerc(table1)
colPerc(table1)
```

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The odd ratio is 0.5 to 1 drugusers community are half as likely to get infected with Covid-19 than the ones who got the placebo. the 95% confidence interval is less than 1 so we know that they are protected. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
In conclusion, the vaccine was effective solely in the women and medicinal consumers, as the odd ratio for both of the categories was 0,5 to 1, which is less than a single, i.e., and it was shown that the medication was effective only in both categories. However, whether you were a male or a homosexual and were taking the vaccine, it is more likely that you would have Covid-19 than the placebo consumer who suggests the vaccine was not effective.